---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PanelReg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
   - \usepackage{amsmath}
   - \usepackage{amssymb}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 0. Setup 

To install the package from Github:

```{r echo = TRUE, eval = FALSE, message=FALSE}
devtools::install_github("setzler/PanelReg")
```

To use the package after it is installed:

```{r echo = TRUE, eval = TRUE, message=FALSE}
library(PanelReg)
```

Make sure these packages have been installed as well:

```{r echo = TRUE, eval = TRUE, message=FALSE}
library(data.table)
library(fixest)
```

### 1. Models included in this package

This package simulates and estimates a class of panel regression models of the following form:

$$
y_{it} = x_{it}'\beta + w_{it}'\gamma + \epsilon_{it} + \nu_{it},
$$

where the data is described as follows:

- $i$ indexes the cross-sectional unit;
- $t$ indexes the time period;
- $y_{it}$ is the outcome of interest;
- $x_{it}$ is the vector of endogenous regressors; and,
- $w_{it}$ is the vector of exogenous regressors

The goal is to estimate $\beta$, while $\gamma$ is treated as a nuisance parameter.

There are three unobservable processes: 

- $\epsilon_{it}$ is the error term upon which $x_{it}$ depends. It is independent of $w_{it}$; 
- $\nu_{it}$ is a serially-independent shock that is independent of $x_{it}, x_{it-1},...$ and $w_{it}, w_{it-1},...$; and,
- $\eta_{it}$ is another serially-independent shock that is independent of $x_{it-1},x_{it-2},...$ and $w_{it}, w_{it-1},...$

The "model" refers to the assumed time-series process governing $\epsilon_{it}$:

1. Fixed Effects ("FE"): $\epsilon_{it} = \bar\epsilon_i, \forall t$.
2. Serially Independent ("iid"): $\epsilon_{it} = \eta_{it}$.
3. Moving Average of Order 1 ("MA1"): $\epsilon_{it} = \eta_{it} + \theta \eta_{it-1}$.
4. First Order Autoregressive  ("AR1"): $\epsilon_{it} = \rho \epsilon_{it-1} + \eta_{it}$.

All of the estimators assume that the number of time periods is small and the sample size is large ("small-T big-N" asymptotics).


### 2. Preliminary components

This package contains only two functions:

- `PanelRegSim`: Simulate panel data according to a model specification.
- `PanelReg`: Estimate $\beta$ given an assumed model.

In order to use PanelReg, we must first define the variable names. The variable names are:

- `id_name`: The name of the variable that indexes the cross-sectional unit.
- `time_name`: The name of the variable that indexes the time period.
- `outcome_name`: The name of the outcome variable.
- `endogenous_names`: A vector of the names of the endogenous regressors.
- `covariate_names`: A vector of the names of the covariates that should be controlled linearly. This argument is optional.
- `fixedeffect_names`: A vector of the names of the covariates that should be controlled by forming discrete group indicators and estimating fixed-effects. Optional.

The following code block defines the variable names for the examples below:

```{r echo=T, eval=T, message=FALSE}
varnames = list(
  id_name = "unit_id",
  time_name = "time_id",
  outcome_name = "outcome",
  endogenous_names = c("endog_var1", "endog_var2"),
  covariate_names = NULL,
  fixedeffect_names = NULL
)
```




### 3. The "FE" model

The error term is given by $\epsilon_{it} = \bar\epsilon_i, \forall t$. The data can be simulated as follows:


```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "FE", seed = 123, sample_size = 30000, 
                             min_year = 2003, max_year = 2006, true_beta = c(0.5, -0.2))
head(simulated_data)
```

Ignoring the endogeneity, the biased estimates are as follows:


```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "exogenous", varnames = varnames)
head(reg)
```


The first-difference of the model is,

$$
y_{it} - y_{it-1} = (x_{it} - x_{it-1})'\beta + (w_{it} - w_{it-1})'\gamma + (\nu_{it} - \nu_{it-1}),
$$

Since the only unobservable in the differenced equation, $\nu_{it} - \nu_{it-1}$, is independent of the regressors, $\beta$ can be estimated without an instrument. The implementation is as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "FE", varnames = varnames)
head(reg)
```




### 4. The "iid" model

The error term is given by $\epsilon_{it} = \eta_{it}$. The model can be simulated as follows:

```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "iid", seed = 123, sample_size = 30000, 
                             min_year = 2003, max_year = 2006, true_beta = c(0.5, -0.2))
head(simulated_data)
```



Ignoring the endogeneity, the biased estimates are as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "exogenous", varnames = varnames)
head(reg)
```


The contemporaneous $x_{it}$ depends on $\eta_{it}$, but its lag $x_{it-1}$ is independent of $\eta_{it}$. Since $x_{it-1}$ is a valid instrument, we can estimate $\beta$ by regressing $y_{it}$ on $x_{it}$, instrumented by $x_{it-1}$, and controlling for $w_{it}$. The implementation is as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "iid", varnames = varnames)
head(reg)
```




### 5. The "MA1" model

The error term is given by $\epsilon_{it} = \eta_{it} + \theta \eta_{it-1}$. The data can be simulated as follows:

```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "MA1", seed = 123, sample_size = 30000, 
                             min_year = 2003, max_year = 2006, true_beta = c(0.5, -0.2))
head(simulated_data)
```




Ignoring the endogeneity, the biased estimates are as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "exogenous", varnames = varnames)
head(reg)
```



Estimating the model as if it were the "iid" model, the biased estimates are as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "iid", varnames = varnames)
head(reg)
```


Similarly to the "iid" case, we can use that lagged values of $x$ are valid instruments for contemporaneous $x$. Since $\epsilon$ is only serially dependent for adjacent periods ($\epsilon_{it}$ is correlated with $\epsilon_{it-1}$ but uncorrelated with $\epsilon_{it-2}$), non-adjacent lags of $x$ are valid instruments. The implementation thus uses $x_{it-2}$ as the instrument. The implementation is as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "MA1", varnames = varnames)
head(reg)
```



### 6. The "AR1" model

The error term is given by $\epsilon_{it} = \rho\epsilon_{it-1} + \eta_{it}$. The data can be simulated as follows: 

```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "AR1", seed = 123, sample_size = 100000, noise_sd = 0, 
                             min_year = 2003, max_year = 2008, true_beta = c(0.5, -0.2))
head(simulated_data)
```





Ignoring the endogeneity, the biased estimates are as follows:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "exogenous", varnames = varnames)
head(reg)
```



The package provides three strategies for estimating the AR(1) model: Panel IV and GMM.

**Panel IV:** The error structure implies that, $$ y_{it} = \rho y_{it-1} + x_{it}'\beta +  x_{it-1}'(-\beta\rho) +  w_{it}'\gamma + w_{it-1}'(-\gamma\rho) + \eta_{it}$$ Since only $x_{it}$ depends on the unobservable $\eta_{it}$, and lagged values of $x$ are independent of contemporaneous $\eta$, it follows that $\beta$ can be estimated by a regression of $y_{it}$ on $x_{it}$, with instrument $x_{it-2}$, controlling for $(y_{it-1}, x_{it-1}, w_{it}, w_{it-1})$. Another available instrument is $y_{it-2}$, so there is an option to use the lagged outcome as an IV as well. 


**GMM:** The error structure implies that, for any guess of $(\beta,\gamma,\rho)$, we can construct the implied $\eta$: $$\hat\eta(\beta,\gamma,\rho) \equiv (y_{it} - \rho y_{it-1}) - (x_{it} - \rho x_{it-1})'\beta - (w_{it} - \rho w_{it-1})'\gamma$$ The GMM approach uses that, at the true values of the parameters, the following vector of moments should hold: $$\mathbb{E}[(x_{it-1},x_{it-2},w_{it}) \hat\eta(\beta,\gamma,\rho) ] = \vec{0}$$ Similar moments should hold with respect to lagged values of the outcome, $y_{it-1}$ and $y_{it-2}$ The implementation is as follows:

**Implementations:** 

Estimating the AR(1) model first requires defining the AR1_options, which is a list of options for the estimation. The options are:

- `AR1_method`: The estimation method to use. The options are "PanelIV" and "GMM". This is required.
- `AR1_IV_outcome`: A logical indicating whether the lagged outcome should be used as an instrument. This is required.
- `AR1_IV_lags`: The number of lags of the outcome to use as instruments. This is required if AR1_method = "GMM". However, it is always 2 for AR1_method = "PanelIV".

Here is the PanelIV implementation if including lagged outcomes as IVs:

```{r echo=T, eval=T, message=FALSE}
AR1_options = list(
  AR1_method = "PanelIV",
  AR1_IV_outcome = TRUE
)

reg = PanelReg(panel_data = simulated_data, panel_model = "AR1", 
               varnames = varnames, AR1_options = AR1_options)
head(reg)
```

Here is the GMM implementation, if using 2 lags and including lagged outcomes as IVs:

```{r echo=T, eval=T, message=FALSE}
AR1_options = list(
  AR1_method = "GMM",
  AR1_IV_outcome = TRUE,
  AR1_IV_lags = 2
)

reg = PanelReg(panel_data = simulated_data, panel_model = "AR1", 
               varnames = varnames, AR1_options = AR1_options)
head(reg)
```

Note that an estimate of the AR(1) persistence parameter, $\rho$, is returned by both "AR1" implementations.



### 7. Controlling for covariates

Consider again the "MA1" model. We can add covariates to the simulated model by setting the `true_gamma` parameter:

```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "MA1", seed = 123, sample_size = 50000, 
                             noise_sd = 0, min_year = 2003, max_year = 2009, 
                             true_beta = c(0.5, -0.2), true_gamma = c(1, -1))

head(simulated_data)
```


If we do not control for the covariates, the estimation becomes biased:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "MA1", varnames = varnames)
head(reg)
```

To control linearly for covariates, we add their names to the varnames list in the `$covariate_names` argument:


```{r echo=T, eval=T, message=FALSE}
varnames$covariate_names = c("co_var1")

reg = PanelReg(panel_data = simulated_data, panel_model = "MA1", varnames = varnames)
head(reg)

varnames$covariate_names = NULL
```


### 8. Controlling for group fixed-effects 

Consider again the "iid" model. We can assign units to groups in the simulated model by setting the `true_FE_var` parameter. When the variance is positive, each unit ID is assigned to a group, and a group-by-time shock is drawn.

```{r echo=T, eval=T, message=FALSE}
simulated_data = PanelRegSim(panel_model = "MA1", seed = 123, sample_size = 50000, 
                             noise_sd = 0, min_year = 2003, max_year = 2008, 
                             true_beta = c(0.5, -0.2), true_FE_var = 5)
head(simulated_data)
```

If we do not control for the group identifiers, the estimation becomes biased by group fixed-effects:

```{r echo=T, eval=T, message=FALSE}
reg = PanelReg(panel_data = simulated_data, panel_model = "MA1", varnames = varnames)
head(reg)
```

To control for group fixed-effects, we add their names to the varnames list in the `$fixedeffect_names` argument. Note that we need to first define a group-by-time identifier:

```{r echo=T, eval=T, message=FALSE}
simulated_data[, group_year := .GRP, by = c("group_id", "time_id")]
varnames$fixedeffect_names = "group_year"

reg = PanelReg(panel_data = simulated_data, panel_model = "MA1", varnames = varnames)
head(reg)

varnames$fixedeffect_names = NULL
```

